---
title: "HW 02 - Basics of Data Frames"
author: "Josh Asuncion"
date: "September 30, 2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 2) Import the data in R

```{r}
dat <- read.csv('data/nba2017-player-statistics.csv', colClasses = c('character', 'character', 'factor', 'character', 'double', rep('integer', 19)))
str(dat)
```

```{r}
library(readr)
dat2 <- read_csv('data/nba2017-player-statistics.csv', col_types = cols(
  .default = col_integer(),
  Player = col_character(),
  Team = col_character(),
  Experience = col_character(),
  Position = col_factor(levels = c('C', 'PF', 'PG', 'SF', 'SG')),
  Salary = col_double()
  )
)
str(dat2)
```

### 3) Right after importing the data

```{r}
dat$Experience <- as.integer(gsub('R', '0', dat$Experience))
```

### 4) Performance of players

```{r}
dat$Missed_FG <- dat$FGA - dat$FGM
dat$Missed_FT <- dat$FTA - dat$FTM
dat$PTS <- (dat$Points3 * 3) + (dat$Points2 * 2) + dat$FTM
dat$REB <- dat$OREB + dat$DREB
dat$MPG <- dat$MIN / dat$GP
dat$EFF <- (dat$PTS + dat$REB + dat$AST + dat$STL + dat$BLK - dat$Missed_FG - dat$Missed_FT - dat$TO) / dat$GP

summary(dat$EFF)
```

```{r}
hist(dat$EFF, main = 'Histogram of Efficiency (EFF)', xlab = 'EFF')
```

```{r}
head(dat[order(dat$EFF, decreasing = TRUE), c('Player', 'Team', 'Salary', 'EFF')], n = 10)
```

```{r}
dat[dat$EFF < 0, 'Player']
```

```{r}
correlations <- data.frame(
  PTS = cor(dat$EFF, dat$PTS),
  REB = cor(dat$EFF, dat$REB),
  STL = cor(dat$EFF, dat$STL),
  AST = cor(dat$EFF, dat$AST),
  BLK = cor(dat$EFF, dat$BLK),
  Missed_FT = -cor(dat$EFF, dat$Missed_FT),
  Missed_FG = -cor(dat$EFF, dat$Missed_FG),
  TO = -cor(dat$EFF, dat$TO)
)
correlations <- correlations[order(correlations$PTS), ]
correlations
```

```{r}
barplot(as.matrix(correlations), main = 'Correlations between Player Stats and EFF', ylim = c(-1,1), cex.names = 0.7)
abline(h = 0)
```

### 5) Efficiency and Salary

```{r}
library(ggplot2)
ggplot(data = dat, aes(x = EFF, y = Salary)) + geom_point() + geom_smooth(method = loess) + labs(x = 'Efficiency') + ggtitle('Scatterplot of Efficiency and Salary')
```

```{r}
cor(dat$EFF, dat$Salary)
```

Based on both the scatterplot of efficiency and salary and on the correlation coefficient between them, it is reasonable to conclude that there exists a positive relationship between efficiency and salary. Mainly, they are related such that an increase in efficiency is correlated with an increase in salary. However, despite the clear correlation, it is not clear if one causes the other. An outside factor could be influencing their relationship.

```{r}
players2 <- dat[dat$MPG >= 20, ]
ggplot(data = players2, aes(x = EFF, y = Salary)) + geom_point() + geom_smooth(method = loess) + labs(x = 'Efficiency') + ggtitle('Scatterplot of Efficiency and Salary')
```

```{r}
cor(players2$EFF, players2$Salary)
```

The relationship between efficiency and salary when only accounting for the more established players is still positive, meaning as efficiency goes up, salary goes up. However, the correlation coefficient is actually smaller than the coefficient when taking into account all players. This is evidenced by the scatterplot, where the data in `players2` is more spread out vertically, causing the loess line to have a less inclined slope. It is reasonable to say that they have a weaker relationship.

### 6) Comments and Reflections

Importing the data was hard, because I didn't have much experience with using either `read.csv` or `read_csv`. I found it was especially difficult using `read_csv` because I couldn't figure out how to change column types.

I found manipulating data frames in Question 4 and 5 to be pretty easy. In addition, I found creating the plots to be less difficult than I thought it would be.

I needed to look at outside resources to figure out how to use `read_csv` and `col_types`, such as the [R Studio blog](https://blog.rstudio.com/2015/04/09/readr-0-1-0/) and the [CRAN website](https://cran.r-project.org/web/packages/readr/README.html). To edit the labels and the titles for the plots using `ggplot`, I needed to use the [R-bloggers website](https://www.r-bloggers.com/how-to-format-your-chart-and-axis-titles-in-ggplot2/).

In all, I spent a few hours on this assignment. I found this homework to be less time consuming than the previous assignment. I had a few problems, such as using `read_csv`, and getting the labels when using `ggplot` to display correctly. However, by the end I was able to fully understand. I felt proud of figuring out how to use the different functions in the `ggplot` library.